---
title: "Leftlet() package"
author: "Chandler, Gabriel, Swarnima, Aiman, Haaken, Terence"
date: "4/14/2019"
output:
  html_document:
    toc: yes
---

# Introduction 
So far in DAVis we have created static maps, which are useful in itself for creating figures for reports and presentations. However, sometimes we may want our audience to interact with the data! This is where `leaflet` comes in - the `leaflet` package overlays your data on top of maps, but also makes those maps interactive.


Leaflet is a popular interactive mapping library written in JavaScript. Leaflet enables you to build the control widget directly while using the map. Every layer you create must be added to the same map, all attributed to a “group“. Then, the control widget will allow you to switch from one group to another.


```{r, echo = FALSE}
#install.packages("leaflet")
library(leaflet)
library(magrittr)
```

Our first example of using `leaflet()` is to map out the best city in the US: Seattle! (This is a very subjective view our dear teammate Chandler holds. Very, very, very subjective). As we learned recently, we will use the pipe operator %>% in the `magrittr` package in order to write our program most effectively. We'll give you a taste of what `leaflet()` can do. While this may not make much sense at the moment, we will explain what both `addTiles()` and `addMarkers()` means later. For now, you just need to know that we are building the images of the map through `addTiles()` and specifying which city is the best city in the US through its longitude and latitude coordinates and labeling it with `addMarkers()`.
```{r}
seattle <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-122.335167, lat=47.608013, popup="The Best City in the US!")
seattle
```

Cool, huh?

# The Map Widget

### `leaflet()` function

The `leaflet()` function returns a Leaflet map widget or interface, which stores a list of objects that can be modified or updated later. The map widget displays an interactive map control. You might have noticed at this point that the `leaflet()` function is not taking any arguments for what data to use. By default, the data argument is set to `NULL`. There are two ways to tell `leaflet()` what data to use. We can either put in values longitude and latitude in the functions that follow, just like we did in the example above where we provide the information about longitude and latitude inside the function `addMarkers()`. The other way is to put a data frame as an argument inside leaflet, and in secondary functions like `addMarkers()`, we can call the latitude and longitue values from the data frame itself. We use a data frame when we have multiple points to plot. We can even use SpatialPolygons. But more on that later.

### Initializing Options
The map widget can be initialized with certain parameters. This is achieved by populating the options argument as shown below.

The most important Map State Options in `leaflet()` are:

1. `crs`: The Coordinate Reference System to use.

2. `center`: The initial geographic center of the map

3. `minZoom`: Minimum zoom level of the map.

4. `maxZoom`: Maximum zoom level of the map.

5. `layers`: Array of layers that will be added to the map initially.

For an extensive list of options outside the scope of this report, please click [here](https://leafletjs.com/reference-1.3.0.html#map-option). 

In this example below, we change the options of the Map Widget by restricting the Maximum Zoom level of the map of the best city in the US. Try changing the value in the `minZoom` and `maxZoom` in order to see the different effect the zoom settings can have on the map. Notice how you can't zoom in more than 8 times.
```{r}
seattle <- leaflet(options = leafletOptions(minZoom = 0, maxZoom = 8)) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-122.335167, lat=47.608013, popup="The Best City in the US!")
seattle
```


### Map Methods
Below are a series of methods to manipulate the map further. 

1. `setView()` sets the center of the map view and the zoom level. In our example below, you can see that the map no longer can zoom in past this starting point. Try changing the values of zoom here in order to fully understand the purpose of `setView()`. 

```{r}
setView(seattle, lng=-122.335167, lat=47.608013, zoom = 10)
```

2. `fitBounds()` fits the view into the rectangle [lng1, lat1] – [lng2, lat2]. `fitBounds()` takes the following parameters: fitBounds(map, lng1, lat1, lng2, lat2, options = list())

3. `clearBounds()` clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided. `clearBounds()` takes the following parameters: clearBounds(map)

# Base Maps

The Base Map is the perfect starting place for learning `leaflet()`. The function of the basemap is to provide fundamental information of the map, acting as a base or a starting point. Then, additional data of specialized nature are compiled or overprinted the base map. The method that we will be focusing on for adding data to the Base Maps is called tiled maps. 

### Tiled Maps

In recent years, as a result of GoogleMaps, the most popular way to display and navigate maps online is to use *tiled maps*. A tiled map is a map that is displayed in a browser that appears to be one large map image, but is really just a number of smaller images, or tiles, drawn next to each other. 

Tiled maps enable everyone to have access to high-quality maps, regardless of their hardware or devices. When displaying a tiled map, only the images covering the geographic area currently being displayed are required. 

One must add tiles to their Base Map in order to get the image of the map to be displayed as the actual geographics you are trying to display. 

In `leaflet()`, the easiest way to add tiles is by calling `addTiles()` with no arguments; by default, OpenStreetMap tiles are used. OpenStreetMap is built by a community of mappers that contribute and maintain data about roads, trails, cafés, railway stations, and much more, all over the world.

```{r}
singapore <- leaflet() %>% setView(lng = 103.8198, lat = 1.3521, zoom = 12)
singapore %>% addTiles()
```

# Markers

Now that we have a great base map, we want to be able to visualize our data here. To do so, we 
use markers to call out points on the map. Marker locations are expressed in latitude/longitude coordinates, and can either appear as icons or as circles.

### Icon Markers
Icon markers are added using the `addMarkers()` or the `addAwesomeMarkers()` functions. We're not kidding. That's a thing.

The default icon is a *dropped pin*. As with most layer functions, the popup argument can be used to add a message to be displayed on click, and the label option can be used to display a text label either on hover or statically.

```{r}
leaflet(data = quakes[1:20,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

In addition to having default icons, you can also make your own icon using the `makeIcon()` function. To make a custom icon, you usually need two images — the actual icon image and the image of its shadow. However, this is out of the scope of this project. 

### Awesome Markers Leaflet Plugin.
Leaflet supports even more customizable markers using the awesome markers leaflet plugin. Told you it was a thing.

The `addAwesomeMarkers()` function is similar to `addMarkers()` function but additionally allows you to specify custom colors for the markers as well as additional icons from numerous different libraries (such as the Font Awesome, Bootstrap Glyphicons, and Ion icons icon libraries).

```{r}
df <- quakes[1:10,]

icons <- awesomeIcons(
  icon = 'close', # the icon argument needs to be the name of any valid icon supported by the the respective library (below two lines)
  iconColor = 'white',
  library = 'ion', # The library argument has to be one of ‘ion’, ‘fa’, or ‘glyphicon’
  markerColor = "red"
)

leaflet(df) %>% addTiles() %>%
  addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(mag))
```

Above, we used an icon from the `ion` library. You can view the extensive list of icons in the ion library  [here](https://ionicons.com/). 

##### Marker Clusters
When there are a large number of markers on a map, you can cluster them using the argument `clusterOptions`. 

In order to make the clusters static, or freeze in their groupings at a specified zoom level, you will need to pass in the argument `freezeAtZoom`. Regardless how far you zoom in, after the variable, or in this instance '5', the clusters become static.

```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions(freezeAtZoom = 5)
)
```

Circle markers are much like regular circles (see [Lines and Shapes]), except that their radius in onscreen pixels stays constant regardless of zoom level.

```{r}
leaflet(quakes) %>% addTiles() %>% addCircleMarkers()

```

# Popups and Labels
Another significant addition Popups are small boxes containing arbitrary HTML, that point to a specific point on the map.

### Pop ups
Use the `addPopups()` function to add a single, standalone popup to your given map.

```{r}
content <- "This is where the largest quake happended in the quakes dataset!"
leaflet(options = leafletOptions(minZoom = 0, maxZoom = 5)) %>% addTiles() %>%
  addPopups(167.62, -15.56, content,
    options = popupOptions(closeButton = FALSE)
  )
```

Popups are typically used when the marker or shape is clicked.  Marker and shape functions in the Leaflet package take a popup argument, where you can pass in HTML to easily attach a simple popup.


```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Long,Lat
Magnitude: 6.4, 167.62, -15.56
Magnitude: 6.1, 169.92, -20.70
Magnitude: 6.0, 65.96, -13.64"
))

leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, popup = ~htmlEscape(Name))
```

The last argument in the above example, `htmltools::htmlEscape` was used to sanitize any characters in the name that might be interpreted as HTML. While we did not need it here, it is important to incldue this argument in order to ensure there is no unnecessary errors due to their being a misinterpretation of the strings. 


### Labels
A label is a HTML comment on markers and shapes that is either always displayed or displayed on when the mouse is held over the marker. Unlike popups, you don’t need to click a marker/polygon for the label to be shown.

You can also customize marker labels using the `labelOptions` argument of the `addMarkers()` function. The `labelOptions` argument can be populated using the `labelOptions()` function. View ?labelOptions for a comprehensive list of possible options for labeling elements. 

```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Long,Lat
Magnitude: 6.4, 167.62, -15.56
Magnitude: 6.1, 169.92, -20.70
Magnitude: 6.0, 65.96, -13.64"
))

leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat,
             label = "6.0 > magnitude",
             labelOptions = labelOptions(noHide = T, textsize = "15px")) %>%
  addMarkers()

```

Finally, one can also add labels without the accompanying markers using the `addLabelOnlyMarkers()` function. We'll go a little more in depth with them later on, when we discuss adding labels to polygons. See [Polygons].

# Color

An important part of any type of data visualization in both static and interactive maps is colors. Colors enable us to highlight different trends, identify distinct variables, and even help convey our findings more directly. While we have already studied colors in DAVis, there is a useful tool that can help you generate your own perfect color palette.

### Pal

`pal` is a function that we ourselves write, which takes a variable and maps that variable to a specific gradient of colour. So say we were plotting Life Expectancy. `pal` would take Life Expectancy as its argument and give each value of Life Expectancy for every country a specific colour.

The function will return the corresponding colors in Hexadecimal. However, it is important to note that the following examples and exercises are able to pass in different arguments into the colors function in the following regards: a palette from the RColorBrewer package, a character vector of RGB or named colors, or the full name of a viridis palette.

##### colorNumeric

`pal` can take both continuous and discrete colours, depending on how we write it. I don't believe we've ever done a map with continuous colours, but if we wanted to, our `pal` would use the function `colorNumeric()`. `colorNumeric()` takes a palette (for example one of those we know from RColorBrewer) and a domain, which is basically the range of values `pal` will be able to accept. Our domain will usually just be the variable we're measuring and colouring by. So say we wanted to colour countries by gdp, our `pal` would look something like this:

```
pal <- colorNumeric(palette = "Greens", domain = df$life_expectancy)
```

##### colorBin
Most of the time in this class however, we've been using discrete colours for a continuous input. For example, we'd have a `breaks` arguments, which would give values and we'd want a different colour for every one of those "bins" between two values. For this we use -- *surprise surprise* -- `colorBin`. `colorBin`, much like `colorNumeric` takes a palette, a domain, but also the number of bins that we want. Note, however, that if you're specifying, there's this annoying default argument called `pretty = TRUE`, which will give your bins a nice rounded breaks, but not necessarily the exact number you wanted. So if you're specifying, it's probably a good idea to set `pretty` to `FALSE`. Overall, should look something like this:

```
binpal <- colorBin("Blues", df$life_expectancy, 6, pretty = FALSE)
```

You can even use your own breaks, like so:

```
breaks <- (0, 10, 30, 50, 100)
binpal <- colorBin("Blues", df$life_expectancy, bins = breaks, pretty = FALSE)
```

The most observant among you will annoyingly go and say: "Wait a minute, what if we want more than 9 bins? Being the excellent, amazing student that I am, I know we can't have more than 9 bins for ColorBrewer's palletes".

Okay, first of all, nobody likes a show-off, don't be obnoxious. Second of all, you'll have to figure that one out on your lonesome during the exercise. Think, though: pal needs another palette, a palette that allows for more colours than 9. If only we'd done something similar in class, developed a function that would allow us to break up a palette into any desired number of colours...

##### Other

By the way, there's also a function called `colorQuantile`, which is basically `colorBin`, but it splits up the input domain into subsets with an equal numbers of values. There's also `colorFactor` for categorical variables. Call me a wizard, but I don't think you'll need those for the exercise.

# Lines and Shapes

The bulk of `leaflet()`. Pay attention now.

### Circles

Adding circles is really not that different from adding markers, especially circle markers, with the small exception that circles have their radii specified in meters. Therefore, circles are scaled to the map, so when the user zooms out, the circles become respectively smaller, while circle markers won't change.

Otherwise, circles take pretty much the exact same arguments as circle markers. To plot them, we use the function `addCircles()`.

Pay attention to the next example, where we use circles to visualize the population of some of the cities in the US. The radii here are a function of the population column from the data frame. 

```{r}

cities <- read.csv(textConnection("
City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994
"))

leaflet(cities) %>% addTiles() %>%
  addCircles(lng = ~Long, lat = ~Lat, weight = 1,
    radius = ~sqrt(Pop) * 30, popup = ~City
  )
```

The above line of code is taken verbatim from the website listed in the sources. We thought it was a good representation and not worth our time to find new long, lat values for new cities.

### Rectangles

Rectangles are a little bit more clunky. As they're rectangles, they need four different long-lat values: two long (lng1 and lng2) and two lat (lat1 and lat2). 

### Polygons

This section is probably the most important section here and also the one we'll be using most extensively, so pay attention.

You should be familiar with `Spatial Polygons` and `SpatialPolygonsDataFrames` (spdf) from class lectures and notes. The former is a function that has long lat values of a polygon (these long lat values can change based on the CRS chosen, as we've already seen in class), whilst the latter are just a data frame of several `Spatial Polygons`.

Well, leaflet uses those too (because leaflet is awesome). This allows us to create amazing chloropleth maps, just like we have done in class, but also with an interactive feature that the maps we've had exposure to in class do not have.

The `addPolygons()` function takes several arguments, the important ones are `weight`, which defines how wide the border (or stroke) should be, `color` is the colour of the border (stroke), `dashArray`, is the line type of the stroke, and `opacity` is how opaque the colour of the stroke should be. So setting `opacity` to 0 will result in no border. Meanwhile, an `opacity` of 1 will mean a white border. `fillOpacity` is similar, meaning how opaque the actual polygon should be. The more complicated ones of note are `fillColor`, `label` and `highlight`.

##### fillColor

The first is fairly self-explanatory. We've done this several times in the past. There is one important and a little confusing difference however. If we wanted the polygons' colour to be the same, then that's simple, we just say `fillColor = "green"`. However, that's boring and competely useless if we have many polygons on the map, especially if we want the colour to represent something. `addPolygons()`'s `fillColor` in that case, takes an argument called `pal`. This argument is so important, we've dedicated an entire section to it in [Color].

##### Labels

Yup, we're still not done with Polygons. Labels are cute and not as complicated. Say you wanted to get information about a specific polygon (or the country it represents) whenever you hovered above it. That's what labels do. For the exercise, there are four things you're going to have to know. You'd want your labels to take in not just "Nevada", but any name you give it. This is done with the replacement that the function `sprintf()` offers. Say you wanted to input muptiple strings without writing them. In that case you'd give `sprintf()` the following: quotes that look like this "%s" and a string or vector of strings that ought to replace it.

```{r}
countryname <- c("Bulgaria", "Canada", "Singapore", "Russia")
labels <- sprintf("%s", countryname)
labels
```

Analogously, if you'd like to input a decimal, you do %d, and for a float: %g. But say you wanted to output this to an HTML widget, which is what you'll have to do if you're using leaflet. Well, all you have to do is add a pipe and pass your label to `lapply(htmltools::HTML)`. The cool thing about passing your text to html is that you can learn a bit of HTML styling. To make text bold, you put `<strong>` before the text and wrap it with `</strong>`. So `<strong>Nevada</strong>` will give you **Nevada**, once in an html widget. `<br/>` means new line in HTML; it's basically like `\n`. Let's see how we can put all that together.

```{r}
countryname <- c("Bulgaria", "Canada", "Singapore", "Russia")
decimal <- c(1L, 3L, 5L, 9L)
float <- c(0.563, 98.3, 4.667, 9.08432487235)

labels <- 
  sprintf("%s is number %d <br/><strong>with</strong> %g", countryname, decimal, float) %>% 
  lapply(htmltools::HTML)

```

The above would output the following, but only in an html widget:

Bulgaria is number 1
**with** 0.563
Canada ...

You can also set labelOptions. To do so, you add the variable to your `addPolygons()`. Several of note are `style`, which uses a list of css (HTML) style additions, which we haven't studied in this course. We'll provide them for the exercise. `textsize`... well, figure it out. `direction` tells the label which way to open. Leaving it to "auto" is fine. Meanwhile, with `labelOptions()` you can also make labels clickable, interactive... it's quite cool, I recommend you check out `?labelOptions()` for more info. You'll only need the above for the exercise though.



##### Highlight

Finally, highlighting. It just makes things look good. When you put your cursor over a specific country, you want that country to be highlighted. We do so by saying `highlight = highlightOptions()` and setting up the variables in `highlightOptions()`. Much like `addPolygons()` itself, `highlightOptions` takes `weight`, `color` and `dashArray` and `fillOpacity`. `fillOpacity` here defines how opaque the highlight should be. So setting `fillOpacity` to 0 will result in an almost completely white cover over the initial polygon. Meanwhile, a `fillOpacity` of 1 will not alter the initial colour at all. It takes all the other standard ones, but these are the ones you'll likely need.

The reason highlight is so cool is because all those variables (weight, fillOpacity etc.) will only materialize over a polygon whenever you move your cursor over that polygon.


# Legends

Adding a legend to your map is another way to increase its usefulness and visual appeal. The process for adding them is much like other legends we have done in the past using other packages.

`addLegend()` takes a `position` argument (like "bottomright", "topleft", etc.), a pal function and the values that that pal function is going to be working with. It can also take `title`, `opacity`, `labels` and others more self-explanatory variables. See `?addLegend()` for more info. 


# Additional Features

While we wish we could have continued to write our report on all the wonderful and awesome features of `leaflet()`, we unfortunately will not be able to cover everything. 

In case you are wanting to explore `leaflet()` further, we wanted to give you a few insights into additional features that can help improve your data analysis and visualization further. 

1. Show/Hide Layers: you can group your data and enable a Interactive Layer Display. When you click the button on the display, it'll populate the map with the corresponding data. 

2. Raster Images: According to Techopedia, Raster graphics, also called bitmap graphics, are digital images that are composed of tiny rectangular pixels, or picture elements, that are arranged in a grid or raster of x and y coordinates in such a way that it forms an image. You can use Raster Images in order to display an additional piece of data, such as looking at the average temperatures in a given month across a specific geographical area while presenting another data variable.

3. Using Shiny with `leaflet()`: Shiny is another R package that hat makes it easy to build interactive web apps.



#Sources

Much of this code is inspired from code present [here](https://rstudio.github.io/leaflet/).
